Guaranteeing Safety and Liveness of Unmanned Aerial Vehicle Platoons on Air Highways

### 2016-01-20 docs/AIAA 2016/high_waypoints.tex
no message

% !TEX root = main.tex
\subsection{From Paths to Waypoints}
Each of the cost-minimizing paths $\hpath_{ij}$ computed from the solution to the Eikonal equation consists of a set of points which are close to adjacent points. To sparsify this set of points to obtain a collection of waypoints, $\wpt_{ij}$, which are also endpoints of air highways, we first add the destination point to the collection and note the path's heading. Next, we add to the collection of waypoints the first point on the path at which the heading changes by some threshold $\theta_C$. This process is repeated along the entire path. Finally, we add the origin points to the collection. We create a collection of these points for all the cost-minimizing paths.

>>>>

% !TEX root = main.tex
\subsection{From Paths to Waypoints}
Each of the cost-minimizing paths $\ppath_i$ computed from the solution to the Eikonal equation consists of a continuous set of points. Each path $\ppath_i$ is an approximation to the sequence of highways $\hws_{N_i}^i = \{\hw^i_j\}_{i=1,j=1}^{i=M,j=N_i}$ defined in \eqref{eq:ahpp}, but now indexes by the corresponding path.

For each path $\ppath_i$, we would like to sparsify the points on the path to obtain a collection of waypoints, $\wpt_{i,j}, j = 1,\ldots, N_i+1$, which are the end points of the highways:

\begin{equation}
\begin{aligned}
\hw^i_j(0) &= \wpt_{i,j}, \\
\hw^i_j(1) &= \wpt_{i,j+1}, \\
j &= 1,\ldots,N_i
\end{aligned}
\end{equation}

 we sparsify this set of points to obtain a collection of waypoints, $\wpt$, which are also endpoints of air highways, we first add the destination point to the collection and note the path's heading. Next, we add to the collection of waypoints the first point on the path at which the heading changes by some threshold $\theta_C$. This process is repeated along the entire path. Finally, we add the origin points to the collection. We create a collection of these points for all the cost-minimizing paths.

### 2016-01-20 docs/AIAA 2016/introduction.tex
no message

% !TEX root = platooning.tex
\section{Introduction}
Unmanned aerial vehicle (UAV) systems have in the past been mainly used for military operations \cite{Tice91}. Recently, however, there has been an immense surge of interest in using UAVs for civil applications through projects such as Amazon Prime Air and Google Project Wing \cite{PrimeAir,ProjectWing,Debusk10}. As a result, government agencies such as the Federal Aviation Administration (FAA) and National Aeronautics and Space Administration (NASA) of the United States are also investigating air traffic control for autonomous vehicles to prevent collisions among potentially numerous UAVs \cite{FAA13}.

Optimal control and game theory are powerful tools for providing liveness and safety guarantees to controlled dynamical systems, and various formulations \cite{Bokanowski10,Mitchell05,Barron89} have been successfully used to analyze problems involving a small number of vehicles \cite{Fisac15,Chen14,Ding08}. These formulations are based on Hamilton-Jacobi (HJ) reachability, which computes the backwards reachable set, defined as the set of states from which a system is guaranteed to have a control strategy to reach a target set of states. Reachability is powerful because it can be used for synthesizing both controllers that steer the system towards a goal (liveness controllers), and controllers that steer the system away from danger (safety controllers). Furthermore, HJ formulations are flexible in terms of system dynamics, enabling the analysis of nonlinear systems. The power and success of HJ reachability analysis in previous applications is evident, especially since numerical tools are readily available to solve the associated HJ Partial Differential Equation (PDE) \cite{LSToolbox,Osher02,Sethian96}. However, the computation is done on a grid, making the problem complexity scale exponentially with the number of states, and therefore with the number of vehicles. This makes the computation intractable for large numbers of vehicles.

A considerable body of work has been done on the platooning of vehicles \cite{Kavathekar11}. For example, \cite{McMahon90} investigated the feasibility of vehicle platooning in terms of tracking errors in the presence of disturbances, taking into account complex nonlinear dynamics of each vehicle.  \cite{Hedrick92} explored several control techniques for performing various platoon maneuvers such as lane changes, merge procedures, and split procedures. In \cite{Lygeros98}, the authors modeled vehicles in platoons as hybrid systems, synthesized safety controllers, and analyzed throughput. Finally, reachability analysis was used in \cite{Alam11} to analyze a platoon of two trucks in order to reduce drag by minimizing the following distance while maintaining collision avoidance safety guarantees.

Previous analyses of a large number of vehicles typically do not provide liveness and safety guarantees to the extent that HJ reachability does; however, HJ reachability typically cannot be used to tractably analyze a large number of vehicles. %In the context of HJ reachability, putting vehicles into platoons is desirable because of the additional structure that platoons impose on its members. With additional structure, pairwise safety guarantees of vehicles can be more easily translated into safety guarantees of all the vehicles in the platoon. 
In this paper, we attempt to reconciliate this trade-off by assuming a single-file platoon, which provides structure that allows pairwise safety guarantees from HJ reachability to translate to safety guarantees for the whole platoon. We first propose a hybrid systems model of UAVs in platoons to capture this structure. Then, we show how HJ reachability can be used to synthesize \textit{liveness controllers} that enable vehicles to reach a set of desired states, and wrap \textit{safety controllers} around the liveness controllers to prevent dangerous configurations such as collisions. Finally, we show simulation results of quadrotors forming a platoon, a platoon responding to a malfunctioning member, and a platoon responding to an outside intruder to illustrate the behavior of vehicles in these scenarios and demonstrate the guarantees provided by HJ reachability.

>>>>

% !TEX root = main.tex
\section{Introduction}
Unmanned aerial vehicle (UAV) systems have in the past been mainly used for military operations \cite{Tice91}. Recently, however, there has been an immense surge of interest in using UAVs for civil applications. Through projects such as Amazon Prime Air and Google Project Wing, companies are looking to send UAVs into the airspace to deliver packages \cite{PrimeAir,ProjectWing}. As a rough estimate, suppose in a city of 2 million people, each person requests a drone delivery every 2 months on average and each delivery requires a 30 minute trip for the drone. This would equate to thousands of vehicles simultaneously in the air. As a result, government agencies such as the Federal Aviation Administration (FAA) and National Aeronautics and Administration (NASA) of the United States are also investigating air traffic control for autonomous vehicles in order to prevent collisions among potentially numerous UAVs \cite{FAA13}. Applications of UAVs extend beyond package delivery; they can also be used, for example, to provide supplies or to firefight in areas that are difficult to reach but require prompt response \cite{Debusk10}.

Optimal control and game theory are powerful tools for providing liveness and safety guarantees to controlled dynamical systems under disturbances, and various formulations \cite{Bokanowski10,Mitchell05,Barron89} have been successfully used to analyze problems involving a small number of vehicles \cite{Fisac15,Chen14,Ding08}. These formulations are based on Hamilton-Jacobi (HJ) reachability, which computes the reachable set, defined as the set of states from which a system is guaranteed to have a control strategy to reach a target set of states. Reachability is a powerful tool because reachable sets can be used for synthesizing both controllers that steer the system towards a set of goal states (liveness controllers), and controllers that steer the system away from a set of unsafe states (safety controllers). Furthermore, the HJ formulations are flexible in terms of system dynamics, enabling the analysis of non-linear systems. The power and success of HJ reachability analysis in previous applications cannot be denied, especially since numerical tools are readily available to solve the associated HJ Partial Differential Equation (PDE) \cite{LSToolbox,Osher02,Sethian96}. However, the computation is done on a grid, making the problem complexity scale exponentially with the number of states, and therefore with the number of vehicles. This makes the computation intractable for large numbers of vehicles.

A considerable body of work has been done on the platooning of vehicles \cite{Kavathekar11}. For example, \cite{McMahon90} investigated the feasibility of vehicle platooning in terms of tracking errors in the presence of disturbances, taking into account complex nonlinear dynamics of each vehicle.  \cite{Hedrick92} explored several control techniques for performing various platoon maneuvers such as lane changes, merge procedures, and split procedures. In \cite{Lygeros98}, the authors modeled vehicles in platoons as hybrid systems, synthesized safety controllers, and analyzed throughput. Finally, reachability analysis was used in \cite{Alam11} to analyze a platoon of two trucks in order to minimize drag by minimizing the following distance while maintaining collision avoidance safety guarantees.

Previous analysis of a large number of vehicles typically do not provide liveness and safety guarantees to the extent that HJ reachability does; however, HJ reachability typically cannot be used to tractably analyze a large number of vehicles. %In the context of HJ reachability, putting vehicles into platoons is desirable because of the additional structure that platoons impose on its members. With additional structure, pairwise safety guarantees of vehicles can be more easily translated into safety guarantees of all the vehicles in the platoon. 
In this paper, we attempt to reconciliate this trade-off by assuming a single-file platoon, which provides structure that allows pairwise safety guarantees from HJ reachability to translate to safety guarantees for the whole platoon. We first propose a hybrid systems model of UAVs in platoons to capture this structure. Then, we show how HJ reachability can be used to synthesize \textit{liveness controllers} that enable vehicles to reach a set of desired states, and wrap \textit{safety controllers} around the liveness controllers in order to prevent dangerous configurations such as collisions. Finally, we show simulation results of quadrotors forming a platoon, a platoon responding to a malfunctioning member, and a platoon responding to an outside intruder to illustrate the behavior of vehicles in these scenarios and demonstrate the guarantees provided by HJ reachability.

### 2016-01-20 docs/AIAA 2016/main.tex
no message

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

%\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
\documentclass[letterpaper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                          % paper

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
\overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

\usepackage{amsmath}    % need for sub equations
\usepackage{amsfonts}
\usepackage{graphicx}   % need for figures
\usepackage{subcaption}
\usepackage{epsfig} 
\usepackage{color}
\usepackage[normalem]{ulem}
\usepackage{cancel}
\usepackage{amssymb}
\usepackage{color}
%\usepackage{my_macros}
\usepackage[ruled,vlined,titlenotnumbered]{algorithm2e}

\title{\LARGE \bf
Safe Platooning of Unmanned Aerial Vehicles via Reachability}

\author{Mo Chen, Qie Hu, Casey Mackin, Jaime F. Fisac, and Claire J. Tomlin
\thanks{This work is supported in part by NSF under CPS:ActionWebs (CNS-0931843) and CPS:FORCES (CNS1239166), by NASA under grants NNX12AR18A and UCSCMCA-14-022 (UARC), by ONR under grants N00014-12-1-0609, N000141310341 (Embedded Humans MURI), and MIT\_5710002646 (SMARTS MURI), and by AFOSR under grants UPenn-FA9550-10-1-0567 (CHASE MURI) and the SURE project. The research of M. Chen has received funding from the ``NSERC PGS-D'' Program. The research of J.F. Fisac has received funding from the ``la Caixa" Foundation.}
\thanks{All authors are with the Department of Electrical Engineering and Computer Sciences, University of California, Berkeley. \{mochen72,qiehu,cmackin,jfisac,tomlin\}@berkeley.edu}
}

\begin{document}
\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%
\begin{abstract}
Recently, there has been immense interest in using unmanned aerial vehicles (UAVs) for civilian operations such as package delivery, firefighting, and fast disaster response. As a result, UAV traffic management systems are needed to support potentially thousands of UAVs flying simultaneously in the airspace, in order to ensure their liveness and safety requirements are met. Hamilton-Jacobi (HJ) reachability is a powerful framework for providing conditions under which these requirements can be met, and for synthesizing the optimal controller for meeting them. However, due to the curse of dimensionality, HJ reachability is only tractable for a small number of vehicles if their set of maneuvers is unrestricted. In this paper, we define a platoon to be a group of UAVs in a single-file formation. We model each vehicle as a hybrid system with modes corresponding to its role in the platoon, and specify the set of allowed maneuvers in each mode to make the analysis tractable. We propose several liveness controllers based on HJ reachability, and wrap a safety controller, also based on HJ reachability, around the liveness controllers. For a single altitude range, our approach guarantees safety for one safety breach; in the unlikely event of multiple safety breaches, safety can be guaranteed over multiple altitude ranges. We demonstrate the satisfaction of liveness and safety requirements through simulations of three common scenarios.
\end{abstract}

\input{introduction.tex}
% Introduction (1p)
%% Motivation
%% Related work
%% Summary of results

\input{formulation.tex}
% Problem formulation (1.5p)
%% A number of quadrotors forming a platoon in a single file
%% Platoon definition and functions

\input{solution.tex}
% Solution methodology (1.5-2p)
%% Hybrid model of vehicles
%% Reachability
%% Functions of platoons: following, joining, splitting

% \input{analysis.tex}
%% Information pattern, stability (0.5p)

\input{results.tex}
% Numerical Simuations (1-1.5p)
%% 1: normal operation: form platoon by joining --> just follow --> 1 quadrotor leaves
%% - Mo
%% 2: vehicles already in platoon, one of them goes rogue
%% - Casey
%% 3: vehicles in platoon, intruder vehicle
%% - Qie
%% 4: vehicles in platoon, another vehicle joining, one vehicle in platoon goes rogue
%%
%% followPlatoon - Jaime

\input{conclusion.tex}

% Conclusion (0.5p)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\addtolength{\textheight}{1cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

 \bibliographystyle{IEEEtran}
 \bibliography{references}
\end{document}

>>>>

\documentclass[journal]{aiaa-pretty}

\newcommand{\pos}{p} % position
\newcommand{\vel}{v} % velocity
\newcommand{\hw}{\mathbb{H}} % a single highway
\newcommand{\hws}{\mathbb{S}} % a sequence of highways
\newcommand{\hwd}{\hat{d}}
\newcommand{\cost}[1]{c_\text{#1}}
\newcommand{\wpt}{\mathcal{W}}
\newcommand{\cmap}{c}
\newcommand{\ccost}{C}
\newcommand{\ppath}{\mathbb{P}}
\newcommand{\ocost}{V}

\author{Mo Chen}
\title{Guaranteeing Safety and Liveness of Unmanned Aerial Vehicle Platoons on Air Highways}

\abstract{Recently, there has been immense interest in using unmanned aerial vehicles (UAVs) for civilian operations such as package delivery, firefighting, and fast disaster response. As a result, UAV traffic management systems are needed to support potentially thousands of UAVs flying simultaneously in the airspace, in order to ensure their liveness and safety requirements are met. Hamilton-Jacobi (HJ) reachability is a powerful framework for providing conditions under which these requirements can be met, and for synthesizing the optimal controller for meeting them. However, due to the curse of dimensionality, HJ reachability is only tractable for a small number of vehicles if their set of maneuvers is unrestricted. In this paper, we define a platoon to be a group of UAVs in a single-file formation. We model each vehicle as a hybrid system with modes corresponding to its role in the platoon, and specify the set of allowed maneuvers in each mode to make the analysis tractable. We propose several liveness controllers based on HJ reachability, and wrap a safety controller, also based on HJ reachability, around the liveness controllers. For a single altitude range, our approach guarantees safety for one safety breach; in the unlikely event of multiple safety breaches, safety can be guaranteed over multiple altitude ranges. We demonstrate the satisfaction of liveness and safety requirements through simulations of three common scenarios.}

\begin{document}

\maketitle

\section*{Nomenclature}
(Nomenclature entries should have the units identified)\\
\noindent\begin{tabular}{@{}lcl@{}}
$x$ &=& System state \\
$\pos=(\pos_x, \pos_y)$ &$=$& Horizontal position \\
$\vel=(\vel_x, \vel_y)$ &$=$& Horizontal velocity
\end{tabular} \\

\input{introduction.tex}
\input{high_form.tex}
\input{high_eikon.tex}
\input{high_waypoints.tex}
\input{high_sim.tex}
\input{plat_form.tex}
\input{plat_reach.tex}
\input{plat_intra.tex}
\input{plat_inter.tex}
\input{plat_safety.tex}
\input{plat_sim.tex}
\input{conclusions.tex}

\section*{Appendix}

\section*{Acknowledgments}
This work is supported in part by NSF under CPS:ActionWebs (CNS-0931843) and CPS:FORCES (CNS1239166), by NASA under grants NNX12AR18A and UCSCMCA-14-022 (UARC), by ONR under grants N00014-12-1-0609, N000141310341 (Embedded Humans MURI), and MIT\_5710002646 (SMARTS MURI), and by AFOSR under grants UPenn-FA9550-10-1-0567 (CHASE MURI) and the SURE project.

\bibliography{references}
\end{document}

### 2016-01-20 docs/AIAA 2016/plat_form.tex
no message

\begin{equation} \label{eq:dyn}
\begin{aligned}
\dot{\pos}_x &= \vel_x, \qquad \dot{\pos}_y = \vel_y  \\
\dot{\vel}_x &= u_x, \qquad \dot{\vel}_y = u_y, \qquad |u_x|,|u_y| \le u_\text{max}
\end{aligned}
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{"fig/vehicleModes"}
	\caption{Hybrid modes for vehicles in platoons.}
	\label{fig:vehicleModes}
\end{figure}

@@

The answers to these questions can be broken down into the maneuvers listed in Section \ref{subsec:platoon_def}. In general, the control strategies of each quadrotor have a liveness component, which specifies a set of states towards which the quadrotor aims to reach, and a safety component, which specifies a set of states that it must avoid. In this paper, we address both the liveness and safety component using reachability analysis.

>>>>

\begin{equation} \label{eq:dyn}
\begin{aligned}
\dot{\pos}_x &= \vel_x \\
\dot{\pos}_y &= \vel_y  \\
\dot{\vel}_x &= u_x \\
\dot{\vel}_y &= u_y \\
|u_x|,|u_y| &\le u_\text{max}
\end{aligned}
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{"fig/vehicleModes"}
	\caption{Hybrid modes for vehicles in platoons.}
	\label{fig:vehicleModes}
\end{figure}

@@

The answers to these questions can be broken down into the maneuvers listed in Section \ref{subsec:platoon_def}. In general, the control strategies of each vehicle have a liveness component, which specifies a set of states towards which the vehicle aims to reach, and a safety component, which specifies a set of states that it must avoid. Together, the liveness and safety controllers guarantee the success and safety of a vehicle in the airspace making any desired mode transition. In this paper, these guarantees are provided using reachability analysis.
